### 2020-12-13 tex/gitlinks.tex
set randomizer seed for test, duh

\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/1b565f496cf77b50c5eb4b5279cee8f687f36b27/tex/figures/#1.py}{\codeicon}\,\,}
\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/1b565f496cf77b50c5eb4b5279cee8f687f36b27/tex/figures/#1.gif}{\animicon}\,\,}
\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/1b565f496cf77b50c5eb4b5279cee8f687f36b27/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
\newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}